---
title: "Yara Digital Ag Solutions – FarmX Case Study"
author: "Doc Anthony"
date: '2022-07-31'
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}
library(dplyr)
library(ggplot2)
library(stringr)
library(gtsummary)

```

## Load and Assess the Data

```{r}

df <- read.csv("CropProductionStatistics.csv")

```

Let's view the shape of the data as well as its data types

```{r}
glimpse(df)

```

Let's see the number of unique values in each variable.

```{r}

sapply(df[1:5], \(x) {
  unique(x) %>% length()
})

```

Are there any missing values?

```{r}

sapply(df, \(x) sum(is.na(x)))

```

The values in season?

```{r}

unique(df$Season)

```

Are there productions every season?

```{r}

df %>% group_by(Season) %>% summarise(missing = sum(is.na(Production)))

df %>% na.omit() %>% 
  group_by(Season) %>% 
  summarise(missing = sum(is.na(Production)),
            zero_prod = sum(Production == 0),
            n = n()
            )

```


## Clean the Data


- Remove the 3730 rows (1.51%) with missing `Production` value.
- Remove the 3523 rows (1.43%) where there was no Production i.e. zero (0).
- Trim the white spaces in `Season`.

```{r}

# remove the rows with missing Production
# trim the extra spaces in Season
# Convert chars to factors
df_clean <- df %>% filter(!is.na(Production), Production != 0) %>% 
  mutate(
    Season = str_trim(Season)
  )

```



## State (Region) and Crop Production


```{r}

top_10 <- df_clean %>% 
  # filter(Season != "Whole Year") %>%
  group_by(State_Name) %>%
  summarise(
    median = median(Production)
  ) %>% 
  arrange(desc(median)) %>% 
  top_n(10) 

top_10 %>%
  ggplot(aes(reorder(State_Name, median), median)) +
  geom_col(fill = "#0C7BDC") +
  coord_flip() +
  labs(title = "Ten States with Highest Production", 
       y = "Average Production (Median)", x = "")


```


Insight: 10 states with the highest crop production are Goa, Haryana, Gujarat, Punjab, Maharashtra, Telangana, Kerata, Andhra Pradesh, Jharkhand and Tamil Nadu.


### Important Districts Among the Top 10 States

Insight: On average, Production is highest in NAWANSHAHR District in Punjab State.

```{r}

df_clean %>% 
  filter(State_Name %in% top_10$State_Name) %>% 
  group_by(District_Name) %>%
  summarise(
    median = median(Production)
  ) %>%
  top_n(10) %>%
  ggplot(aes(reorder(District_Name, median), median)) +
  geom_col(fill = "#0C7BDC") +
  labs(title = "In the Top 10 States, Production is highest in Winter", 
       y = "Average Production (Median)", x = "") +
  coord_flip()
```


### Season and Crop Production

```{r}

df_clean %>% select(Season, Area, Production) %>%
  mutate(
    Production_per_area = round(Production / Area, 2) 
  ) %>%
  select(-Area) %>%
  tbl_summary(
    by = "Season",
    type = all_continuous() ~ "continuous2",
    statistic = all_continuous() ~ c("{mean} ({sd})",
                                     "{median} ({p25}, {p75})", 
                                     "{min}, {max}")
    ) %>%
  add_p()

```

Insight: Production is highest in Winter.

```{r}

df_clean %>% 
  # filter(Season != "Whole Year") %>% 
  group_by(Season) %>%
  summarise(
    median = median(Production)
  ) %>%
  ungroup() %>%
  ggplot(aes(reorder(Season, median), median)) +
  # use color-blind friendly color
  geom_col(fill = "#0C7BDC") +
  labs(title = "On Average, Production is highest in Winter", 
       y = "Average Production (Median)", x = "Season")
```

Among the top 10 states, production is still highest in winter

```{r}

df_clean %>% 
  filter(State_Name %in% top_10$State_Name) %>% 
  group_by(Season) %>%
  summarise(
    median = median(Production)
  ) %>%
  ungroup() %>%
  ggplot(aes(reorder(Season, median), median)) +
  # use color-blind friendly color
  geom_col(fill = "#0C7BDC") +
  labs(title = "In the Top 10 States, Production is highest in Winter", 
       y = "Average Production (Median)", x = "Season")
```


## Crops and Production

Insight: The Top 5 crops are coconut, wheat, total foodgrain, sugar and rubber.

```{r}

df_clean %>% 
  group_by(Crop) %>%
  summarise(
    median = median(Production)
  ) %>%
  top_n(10) %>%
  ggplot(aes(reorder(Crop, median), median)) +
  geom_col(fill = "#0C7BDC") +
  labs(title = "The top 10 crops", 
       y = "Average Production (Median)", x = "") +
  coord_flip()

```


### 10 5 crops by State and Season


```{r}

top_5_crops <- df_clean %>% 
  group_by(Crop) %>%
  summarise(
    median = median(Production)
  ) %>%
  top_n(5) %>% select(Crop)

top_5_crops_in_top_10_states <-  df_clean %>% 
  filter(State_Name %in% top_10$State_Name, Crop %in% top_5_crops$Crop)
  
top_5_crops_in_top_10_states %>%
  group_by(State_Name, Crop) %>%
  summarise(
    median = median(Production)
  ) %>%
  top_n(10)

top_5_crops_in_top_10_states %>%
  group_by(Season, Crop) %>%
  summarise(
    median = median(Production)
  ) %>%
  top_n(10)

```


## Land Area and Production

Insight: Production Increases with Land Area

```{r}

df_clean %>%
  filter(State_Name %in% top_10$State_Name) %>% 
  ggplot(aes(log(Area), log(Production))) +
  geom_point() +
  geom_jitter(alpha = 0.5)
```


### Land Area and Production with Respect to the Top 5 Crops in the Top 10 States/Region

```{r}

top_5_crops_in_top_10_states %>%
  ggplot(aes(log(Area), log(Production), color = Crop)) +
  geom_point() +
  geom_jitter(alpha = 0.5)

```

